---
title: "Monte Carlo Methods and Computer Arithmetics"
subtitle: "HW 2 of STAT 5361 Statistical Computing"
author: Biju Wang^[<bijuwang@uconn.edu>]
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
header-includes: 
  \usepackage{float}
  \floatplacement{figure}{H}
output: 
  pdf_document:
    number_sections: true
    
abstract: Use Monte Carlo Methods to approch $\Phi(t)$ and explain some computer arithmetics.
keywords: Monte Carlo, Arithmetics
---

# Monte Carlo Methods
## Principles
The CDF of standard norm distribution is
\begin{equation}
\Phi(t) = \int^{t}_{-\infty}\frac{1}{\sqrt{2\pi}}e^{-y^{2}/2}dy
\end{equation}
by the Monte Carlo methods
\begin{equation}
\hat{\Phi}(t)=\frac{1}{n}\sum^{n}_{i=1}I(X_{i}\leq t)
\end{equation}
where $X_{i}$'s are iid $N(0, 1)$ variables.

## Approximation Outcomes
The approximation is implemented at $n\in \{10^{2}, 10^{3}, 10^{4}\}$ at $t\in\{0.0, 0.67, 0.84, 1.28, 1.65, 2.32, 2.58, 3.09, 3.72\}$. The outcome table with true values is shown below.

$\hat{\Phi}(t)$|$t=0.0$|$t=0.67$|$t=0.84$|$t=1.28$|$t=1.65$|$t=2.32$|$t=2.58$|$t=3.09$|$t=3.72$|
:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|
$n=10^{2}$|0.4600|0.7400|0.8200|0.9100|0.9700|0.9900|1.0000|1.0000|1.0000|
$n=10^{3}$|0.5210|0.7390|0.7940|0.8920|0.9360|0.9900|0.9960|0.9990|0.9990|
$n=10^{4}$|0.5056|0.7496|0.7937|0.8967|0.9521|0.9910|0.9960|0.9994|1.0000|
$\Phi(t)$ |0.5000|0.7486|0.7995|0.8997|0.9505|0.9898|0.9951|0.9990|0.9999|
Table: Approximation Outcomes with True Values

```{r}
n <- c(10^2, 10^3, 10^4)
t <- c(0.0, 0.67, 0.84, 1.28, 1.65, 2.32, 2.58, 3.09, 3.72)
truth <- pnorm(t)

table <- matrix(NA, 3, 9)

set.seed(1)
sample <- list()
for (i in 1:length(n)) {
  sample[[i]] <- rnorm(n[i])
}

for (i in 1:length(n)) {
  for (j in 1:length(t)) {
    table[i, j] <- sum(sample[[i]] <= t[j])/n[i]
  }
}

```

## Box Plots of the Bias
For each $n$ and $t$, we have $\hat{\Phi}(t)$ and the true value $\Phi(t)$, then we can calculate the bias. We repeat the experiments $100$ times and will obtain $100$ biases. Thus box plots can be drawn.

```{r, echo = FALSE, fig.height = 5.8, fig.cap = "Side-by-Side Boxplots of Bias at all $t$ for Different $n$"}
library("ggplot2")
library("reshape")
library("gridExtra")

bias1 <- matrix(NA, 100, 9)
bias2 <- matrix(NA, 100, 9)
bias3 <- matrix(NA, 100, 9)

for (i in 1:100) {
  set.seed(i)
  for (j in 1:length(t)) {
    bias1[i, j] <- sum(rnorm(10^2) <= t[j])/10^2 - truth[j]
    bias2[i, j] <- sum(rnorm(10^3) <= t[j])/10^3 - truth[j]
    bias3[i, j] <- sum(rnorm(10^4) <= t[j])/10^4 - truth[j]
  }
}

bias1 <- as.data.frame(bias1)
bias2 <- as.data.frame(bias2)
bias3 <- as.data.frame(bias3)
colnames(bias1) <- c("t=0.0", "t=0.67", "t=0.84", "t=1.28", "t=1.65", "t=2.32", "t=2.58", "t=3.09", "t=3.72")
colnames(bias2) <- c("t=0.0", "t=0.67", "t=0.84", "t=1.28", "t=1.65", "t=2.32", "t=2.58", "t=3.09", "t=3.72")
colnames(bias3) <- c("t=0.0", "t=0.67", "t=0.84", "t=1.28", "t=1.65", "t=2.32", "t=2.58", "t=3.09", "t=3.72")

bias1_melt <- melt(bias1)
bias2_melt <- melt(bias2)
bias3_melt <- melt(bias3)

p1 <- ggplot(bias1_melt, aes(x = factor(variable), y = value)) + geom_boxplot() + labs(x = expression("Values of"~t), y = "Bias") + theme(plot.title = element_text(hjust = 0.5)) + ggtitle(expression("Side-by-Side Boxplots of Bias at All"~t~"for"~n == 10^2))
p2 <- ggplot(bias2_melt, aes(x = factor(variable), y = value)) + geom_boxplot() + labs(x = expression("Values of"~t), y = "Bias") + theme(plot.title = element_text(hjust = 0.5)) + ggtitle(expression("Side-by-Side Boxplots of Bias at All"~t~"for"~n == 10^3))
p3 <- ggplot(bias3_melt, aes(x = factor(variable), y = value)) + geom_boxplot() + labs(x = expression("Values of"~t), y = "Bias") + theme(plot.title = element_text(hjust = 0.5)) + ggtitle(expression("Side-by-Side Boxplots of Bias at All"~t~"for"~n == 10^4))
grid.arrange(p1, p2, p3, nrow = 3)
```


# Computer Arithmetics

